Race-based disparities in allocation of academic disciplinary actions are associated

### 2018-02-26 written_docs/draft.Rmd
There's some content here.

---
title             : "Race-based disparities in allocation of academic disciplinary actions are associated with county-level rates of bias"
shorttitle        : "Discipline Disparities"

author: 
  - name          : "Travis Riddle"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "triddle@princeton.edu"
  - name          : "Stacey Sinclair"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Princeton University"

author_note: >

  Enter author note here.

abstract: >
  Enter abstract here.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["ref.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_word
---

@@

# Methods
## Data Sources
```{r summary-data}
library(dplyr)
library(forcats)
library(haven)
library(tidyr)
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
                         colClasses = 'character')
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
model.data <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')

@@

```
We used three distinct data sources for the work described here. The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education. The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black[^1]. Previous work using these data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we]. After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000, 1)` million are white and `r  round(enrollment_totals$ns[1]/1000000, 1)` million are black.

@@

df %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age)  -> subdat

df2 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

@@

```{r teachers-summary}

>>>>

---
title: Race-based disparities in allocation of academic disciplinary actions are associated
  with county-level rates of bias
author:
- address: ''
  affiliation: '1'
  corresponding: yes
  email: triddle@princeton.edu
  name: Travis Riddle
- affiliation: '1'
  name: Stacey Sinclair
affiliation:
- id: '1'
  institution: Princeton University
output:
  papaja::apa6_word: default
  papaja::apa6_pdf: default
bibliography: ref.bib
class: man
figsintext: yes
figurelist: no
footnotelist: no
keywords: keywords
lang: english
lineno: yes
author_note: |
  Enter author note here.
shorttitle: Discipline Disparities
tablelist: no
abstract: |
  Enter abstract here.
wordcount: X
---

@@

# Methods
## Data Sources
```{r summary-data}
library(dplyr)
library(forcats)
library(haven)
library(tidyr)
library(ggplot2)
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
                         colClasses = 'character')
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
model.data <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')

@@

```
We used three distinct data sources for the work described here. The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education. The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black[^1]. Previous work using these data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we]. After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000)` million are white and `r  round(enrollment_totals$ns[1]/1000000)` million are black.

@@

df %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df2 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

@@

Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data. Implicit bias was assessed via an Implicit Association Test (IAT). This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses. The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions. This association is indicated in the IAT D-score, which we used as a measure of implicit bias. Our measure of explicit bias is the difference between reported warmth towards whites (i.e. *how warm or cold do you feel towards Whites? 0=very cold, 10=very warm*) and reported warmth toward blacks.

# Results
## Project Implicit Estimates
```{r exp-diff}
individual_data %>%
  mutate(explicit_bias_diff = twhite_0to10-tblack_0to10) -> individual_data
```
We first report the results of estimating the implicit and explicit biases in from Project Implicit data. Overall, the individuals in project implicit show a pro-white bias in both implicit (*mean* = `r round(mean(individual_data$D_biep.White_Good_all), 2)`, *sd* = `r round(sd(individual_data$D_biep.White_Good_all), 2)`), and explicit measures (*mean* = `r round(mean(individual_data$explicit_bias_diff, na.rm=T), 2)` *sd* = `r round(sd(individual_data$explicit_bias_diff, na.rm=T), 2)`). When aggregated at the county level and adjusted with poststratification, the summary statistics across counties are similar in terms of their location, but as expected, the variability is much diminished ($mean_{implicit}$ = `r round(mean(as.numeric(county_means$weighted_bias), na.rm=T), 2)`, $sd_{implicit}$ = `r round(sd(as.numeric(county_means$weighted_bias), na.rm=T), 2)`; $mean_{explicit}$ = `r round(mean(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, $sd_{explicit}$ = `r round(sd(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, where on both scales 0 = no bias, and positive numbers indicate a pro-white bias).

## Disciplinary action frequency

```{r disc-count, results='asis'}
model.data %>% 
  filter(exclude==FALSE) %>%
  group_by(group, metric) %>% 
  summarise(students = sum(number)) %>%
  mutate(metric = fct_recode(metric, 
                             'expulsions' = 'expulsion_combined',
                             'school arrests' = 'in_school_arrest',
                             'in-school suspension' = 'inschool_susp',
                             'law enforcement referral' = 'law_enforcement',
                             'out-of-school suspension' = 'oos_susp')) -> counts_table

apa_table(counts_table, caption = 'Count of students by race receiving each type of disciplinary action')
```

Table \@ref(tab:disc-count) shows the number of students of each race who were reported having received each of the actions under consideration. The counts range from a low of just `r counts_table$students[7]` white students arrested to a high of `r counts_table$students[5]` black students receiving an out-of-school suspension. Considering the vast differences in the overall number of black and white students, this simple count already illustrates that black students are disciplined at rates far higher than their white counterparts.

## Associations across county
Figure \@ref(fig:overall-associations) shows the estimate of primary interest for each of the models. The estimates displayed are the coefficients for the interaction between race and each of the two bias measurements. Given that African Americans are the baseline group, negative values for this coefficient indicate that as the bias in a county increases, the gap between the probability of a black student being disciplined and the probability of a white student being disciplined grows.

```{r overall-associations, fig.cap="Association between each metric and county-level estimates of explicit and implicit bias. Negative values indicate that the rate of increase (or decrease) for blacks is faster (or slower) than for whites. Point is the mean of the posterior and error bars represent 95% bayesian uncertainty intervals."}
#the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff/'
plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(14,4000,length(files)))
  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    df <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate) %>%
      filter(!is.na(county_id)) -> raw_rate
    posterior_combo[,,j] <- t(df[,c(1:14)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (p in 1:14){
    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
  }

  data.frame(est=mean(p_cons[13,]),
             lower=quantile(p_cons[13,], .025),
             upper=quantile(p_cons[13,], .975),
             p_neg = prop.table(table(p_cons[13,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3$metric <- fct_recode(plot.dat3$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat3_all <-plot.dat3
plot.dat3_all$bias <- 'Implicit'

#weighted warmth w/exclusions & explicit difference
plot.dat <- data.frame(group = NA, weighted_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)

for(i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(14,4000,length(files)))
  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    df <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_warmth')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate_warmth) %>%
      filter(!is.na(county_id)) -> raw_rate_warmth
    posterior_combo[,,j] <- t(df[,c(1:14)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (p in 1:14){
    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
  }

  data.frame(est=mean(p_cons[14,]),
             lower=quantile(p_cons[14,], .025),
             upper=quantile(p_cons[14,], .975),
             p_neg = prop.table(table(p_cons[14,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3$bias <- 'Explicit'
plot.dat3$metric <- fct_recode(plot.dat3$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
  geom_point(position=position_dodge(width=.5), size=2) +
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                width=.5, position=position_dodge(width=.5)) +
  theme_classic() +
  coord_flip() +
  geom_hline(yintercept=0) +
  ylab('Black-White Difference in log-odds slope') +
  theme(axis.title.y = element_blank(),
        legend.title = element_blank())
```

Several patterns are apparent from this figure. First, with the exceptions of implicit bias and expulsions and law enforcement referrals, all estimates are directionally consistent with higher levels of bias leading to larger differences between groups. Second, these effects are especially consistent for the two types of suspensions. The largest effect estimated is between implicit bias and out-of-school suspensions. The difference in the slope of the association between implicit bias and the log of the odds for out-of-school suspensions between white and black students is estimated to be `r round(plot.dat3_all$est[1], 2)`, with 95% of the posterior distribution between `r round(plot.dat3_all$lower[1], 2)` and `r round(plot.dat3_all$upper[1], 2)` and a proportion `r afex::round_ps(1)` of the posterior distribution consistent with a negative effect.

Although not nearly as large of a difference, we have similar certainty with respect to the association between out-of-school suspensions and explicit bias. The difference in the slope of the association between explicit bias and the log of the odds for out-of-school suspensions between white and black students is estimated to be `r round(plot.dat3_all$est[6], 2)`, with 95% of the posterior distribution between `r round(plot.dat3_all$lower[6], 2)` and `r round(plot.dat3_all$upper[6], 2)` and a proportion `r afex::round_ps(plot.dat3_all$p_neg[6])` of the posterior distribution consistent with a negative effect.

The estimated associations for in-school suspensions are smaller still, but are generally consistent with effects of the same direction. For implicit bias, the relevant parameter is estimated at `r round(plot.dat3_all$est[3], 2)` [`r round(plot.dat3_all$lower[3], 2)`, `r round(plot.dat3_all$upper[3], 2)`] $p_{neg}$ = `r round(plot.dat3_all$p_neg[3], 2)`, and for explicit bias, the effect is slightly larger, and a positive effect is essentially not credible, given the data and model `r round(plot.dat3_all$est[8], 2)` [`r round(plot.dat3_all$lower[8], 2)`, `r round(plot.dat3_all$upper[8], 2)`] $p_{neg}$ `r afex::round_ps(plot.dat3_all$p_neg[8])`.

Other outcomes are estimated with less precision, or with patterns that are inconsistent between implicit and explicit bias. For instance, examining the associations for expulsions, the effect of explicit biases are generally in the expected direction (*est* = `r round(plot.dat3_all$est[10], 2)`, [`r round(plot.dat3_all$lower[10], 2)`, `r round(plot.dat3_all$upper[10], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[10])`), but the effect for implicit bias is estimated to be close to zero, with a enough uncertainty (*est* = `r round(plot.dat3_all$est[5], 2)`, [`r round(plot.dat3_all$lower[5], 2)`, `r round(plot.dat3_all$upper[5], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[5])`) to make it difficult to claim an effect of one direction or the other. The model indicates similar uncertainty with respect to school-related arrests and both estimates of bias ($est_{explicit}$ = `r round(plot.dat3_all$est[9], 2)`, [`r round(plot.dat3_all$lower[9], 2)`, `r round(plot.dat3_all$upper[9], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[9])`; $est_{implicit}$ = `r round(plot.dat3_all$est[4], 2)`, [`r round(plot.dat3_all$lower[4], 2)`, `r round(plot.dat3_all$upper[4], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[4])`).

```{r detail-fig-exp, fig.cap="Association between bias and expulsions. Top: Association between bias and the estimated probability of expulsion. Line is the mean of the posterior. Bands indicate 95% uncertainty intervals; Bottom: Association between bias and the relative risk ratio for black students to white students. Points represent counties, whose size are scaled to the number of schools in that county."}
library(gridExtra)
i <- 'expulsion_combined'
pth <- paste(base_path, i, '/', sep='')

files <- list.files(pth)
j <- 1
posterior_combo <- array(0, dim=c(14,4000,length(files)))
raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, 
                       weighted_warmth=NA, group=NA, nincidents=NA, 
                       nstudents=NA, rate=NA)
for (k in files){
  #print(i)
  load(paste(pth,k,sep=''))
  df <- as.matrix(m)
  m$data %>% 
    select(county_id, COMBOKEY) %>% 
    distinct() %>% 
    group_by(county_id) %>% 
    summarise(nschools=n()) -> n_schools
  m$data %>%
    group_by(county_id, group) %>%
    summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
    mutate(rate=nincidents/nstudents) -> rate
  n_schools %>%
    left_join(m$data[, c('county_id', 'weighted_bias', 'weighted_warmth')]) %>%
    distinct() %>%
    left_join(rate) %>%
    rbind(raw_rate) %>%
    filter(!is.na(county_id)) -> raw_rate
  posterior_combo[,,j] <- t(df[,c(1:14)])
  j <- j+1
}
p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

for (p in 1:14){
  m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
}
for (p in 1:14){
  m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
}
for (p in 1:14){
  m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
}
for (p in 1:14){
  m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
}

newdat <- expand.grid(group = rownames(table(m$data$group)),
                      total_pop=0,
                      unemp_rate=0,
                      med_income=0,
                      poverty_rate=0,
                      col_grads=0,
                      white_prop=0,
                      black_prop=0,
                      b.w.ratio=0,
                      weighted_bias = seq(-2.5,2.5,.25),
                      weighted_warmth = 0,
                      number=0,
                      total_number=1)
y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

cbind(newdat, t(y_hat)) %>% 
  mutate(index=row_number()) %>%
  gather(sample, value, `1`:`4000`) -> posterior_distribution

posterior_distribution %>%
  select(group, weighted_bias, sample, value, index) %>%
  group_by(group, weighted_bias) %>%
  summarise(est = mean(value),
            lower = quantile(value, .025),
            upper = quantile(value, .975)) %>%
  ungroup() %>%
  mutate(bias = 'Implicit') -> plot.dat

newdat <- expand.grid(group = rownames(table(m$data$group)),
                      total_pop=0,
                      unemp_rate=0,
                      med_income=0,
                      poverty_rate=0,
                      col_grads=0,
                      white_prop=0,
                      black_prop=0,
                      b.w.ratio=0,
                      weighted_bias = 0,
                      weighted_warmth = seq(-2.5,2.5,.25),
                      number=0,
                      total_number=1)
y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

cbind(newdat, t(y_hat)) %>% 
  mutate(index=row_number()) %>%
  gather(sample, value, `1`:`4000`) -> posterior_distribution

posterior_distribution %>%
  select(group, weighted_warmth, sample, value, index) %>%
  mutate(weighted_bias = weighted_warmth) %>%
  group_by(group, weighted_bias) %>%
  summarise(est = mean(value),
            lower = quantile(value, .025),
            upper = quantile(value, .975)) %>%
  ungroup() %>%
  mutate(bias = 'Explicit') %>%
  rbind(plot.dat) -> plot.dat

ggplot(plot.dat, aes(x=weighted_bias, y=est, group=group, linetype=group)) +
  geom_line(aes(color=group)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
  theme_classic() +
  ylab('Probability \n of Expulsion') +
  xlab('Standardized Bias') + 
  facet_wrap(~bias, ncol=2) +
  xlim(-3.5, 3.5) +
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) -> p1

posterior_distribution %>%
  select(group, weighted_warmth, sample, value) %>%
  spread(group, value) %>%
  mutate(ratio = black/white) %>%
  mutate(weighted_bias = weighted_warmth) %>%
  group_by(weighted_bias) %>%
  summarise(est = mean(ratio),
            lower = quantile(ratio, .025),
            upper = quantile(ratio, .975)) %>%
  ungroup() %>%
  mutate(bias = 'Explicit') -> plot.dat

newdat <- expand.grid(group = rownames(table(m$data$group)),
                      total_pop=0,
                      unemp_rate=0,
                      med_income=0,
                      poverty_rate=0,
                      col_grads=0,
                      white_prop=0,
                      black_prop=0,
                      b.w.ratio=0,
                      weighted_bias = seq(-2.5,2.5,.25),
                      weighted_warmth = 0,
                      number=0,
                      total_number=1)
y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

cbind(newdat, t(y_hat)) %>% 
  mutate(index=row_number()) %>%
  gather(sample, value, `1`:`4000`) -> posterior_distribution

posterior_distribution %>%
  select(group, weighted_bias, sample, value) %>%
  spread(group, value) %>%
  mutate(ratio = black/white) %>%
  group_by(weighted_bias) %>%
  summarise(est = mean(ratio),
            lower = quantile(ratio, .025),
            upper = quantile(ratio, .975)) %>%
  ungroup() %>%
  mutate(bias = 'Implicit') %>%
  rbind(plot.dat) -> plot.dat

raw_rate %>%
  mutate(Implicit = weighted_bias,
         Explicit = weighted_warmth) %>%
  select(-weighted_bias, -weighted_warmth) %>%
  gather(bias, value, Implicit:Explicit) %>%
  select(county_id, nschools, bias, group, rate, value) %>%
  spread(group, rate) %>%
  mutate(odds_r = black/white) -> raw_or

ggplot(plot.dat, aes(x=weighted_bias, y=est)) +
  geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.075) +
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
  theme_classic() +
  #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
  scale_alpha_manual(values=c(.2, 1), guide=F) +
  ylim(0,9) +
  xlim(-3.5, 3.5) + 
  xlab('Standardized Bias') +
  ylab('Relative \n Risk Ratio') +
  geom_hline(yintercept=1) +
  facet_wrap(~bias) -> p2

p1$data %>% filter(weighted_bias %in% c(0, 1)) -> rates
p2$data %>% filter(weighted_bias %in% c(0, 1)) -> risks

grid.arrange(p1, p2)
```
